---
title: "whdsm"
author: "wyg"
date: "2020/10/17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## dplyr
```{r}
library(tidyverse)
gapminder <- read_csv("D:/class/data software/data/gapminder.csv")
View(gapminder)
head(gapminder)
tail(gapminder)
head(gapminder, 10)
str(gapminder)
```

+ **filter()**: pick observations by their values
+ **select()**: pick variables by their names
+ **mutate()**: create new variables with functions of existing variables
+ **summarise()**: collapse many values down to a single summary
+ **arrange()**: reorder the rows

These can all be used in conjunction with **group_by()** which changes the scope of each function from operating on the entire dataset to operating on it group-by-group. 

### 1.filter()
```{r}
filter(gapminder, lifeExp < 29)
filter(gapminder, country == "Mexico")
filter(gapminder, country %in% c("Mexico", "Peru"))
filter(gapminder, country == "Mexico", year == 2002)
clexp <-  
  c(filter(gapminder, country == "Brazil", year >= 1987)["lifeExp"])
cl <- as.vector(unlist(clexp)) 
mean(cl)
```


### 2.select()
```{r}
gap1 <- dplyr::select(gapminder, year, country, lifeExp)
head(gap1, 3)
gap2 <- dplyr::select(gapminder, year:lifeExp)
head(gap2, 3)
gap3 <- dplyr::select(gapminder, 1, 2, 4)
head(gap3, 3)
gap4 <- dplyr::select(gapminder, -continent, -lifeExp)
head(gap4, 3)
```


#### pipe %>% operator
```{r}
gap_cambodia <- gapminder %>%
  filter(country == "Cambodia") %>%
  dplyr::select(-continent, -lifeExp)
gap_cambodia
```


### 3.mutate()
```{r}
gapminder %>%
  mutate(gdp = pop * gdpPercap)
gapminder %>%
  filter(continent == "Asia", year == 2007) %>%
  mutate(population = pop / 1000)

```


